A Roadmap for\\HEP Software and Computing R\&D\\for the 2020s

### 2018-11-09 CWP/papers/HSF-CWP-2017-01_roadmap/latex/cwp-main.tex
Add reference for FNAL acceleator schedule
add references (#112)
* add references

* Update cwp-main.tex

add missed reference

\begin{figure*}
\vspace*{0.3cm}
    \centering
    \includegraphics[width=0.94\textwidth]{10yr-PLAN-9Feb-2017-V7a.pdf}
    \caption{Run schedule for the Fermilab facility until 2026.}
    \label{fig:fnal-if}
\end{figure*}

@@

While the density of transistors on silicon continues to increase
following Moore's Law (albeit more slowly than in the past), power
density constraints have limited the clock speed of processors for more
than a decade. This has effectively stalled any progress in the
processing capacity of a single CPU core. Instead, increases in
potential processing capacity come from increases in the core count of
CPUs and wide CPU registers. Alternative processing architectures have
become more commonplace. These range from the many-core architecture
based on standard x86\_64 cores to numerous alternatives such as GPUs.
For GPUs, the processing model is very different, allowing a much
greater fraction of the die to be dedicated to arithmetic calculations,
but at a price in programming difficulty and memory handling for the
developer that tends to be specific to each processor generation.
Further developments may even see the use of FPGAs for more
general-purpose tasks. Fully exploiting these evolutions requires a
shift in programming model to one based on \emph{concurrency}.

@@

Particle physics is no longer alone in facing these massive data
challenges. Experiments in other fields, from astronomy to genomics,
will produce huge amounts of data in the future, and will need to
overcome the same challenges that we face, i.e., massive data handling and
efficient scientific programming. Establishing links with these fields
has already started. Additionally, interest from the computing science
community in solving these data challenges exists, and mutually
beneficial relationships would be possible where there are genuine
research problems that are of academic interest to that community and
provide practical solutions to ours. The efficient processing of massive
data volumes is also a challenge faced by industry, in particular the
internet economy, which developed novel and major new technologies under
the banner of \emph{Big Data} that may be applicable to our use cases
\cite{47224,Shanahan:2015:LSD:2783258.2789993,Armbrust:2015:SSR:2723372.2742797}.

@@

\paragraph{Fast Simulation} The work on Fast Simulation is also
accelerating with the objective of producing a flexible framework that
permits Full and Fast simulation to be combined for different particles
in the same event. Various approaches to Fast Simulation are being tried
all with the same goal of saving computing time, under the assumption
that it is possible to improve time performance without an unacceptable
loss of physics accuracy. There has recently been a great deal of
interest in the use of Machine Learning in Fast Simulation, most of
which has focused on the use of multi-objective regression and
generative adversarial networks (GANs). Since use of GANs allows for
non-parametric learning in cases such as calorimetric shower
fluctuations, it is a promising avenue for generating non-Gaussian and
highly correlated physical effects. This is an obvious area for future
expansion and development, as it is currently in its infancy.

@@

Over the past 20 years the HEP community has developed and gravitated
around a single analysis ecosystem based on ROOT~\cite{Brun1996}.
ROOT is a general-purpose object
oriented framework that addresses the selection, integration,
development, and support of a number of foundation and utility class
libraries that can be used as a basis for developing HEP application
codes. The added value to the HEP community is that it provides an
integrated and validated toolkit, and its use encompasses the full event
processing chain; it has a major impact on the way HEP analysis is
performed. This lowers the hurdle to start an analysis, enabling the
community to communicate using a common analysis language, as well as
making common improvements as additions to the toolkit quickly become
available. The ongoing ROOT programme of work addresses important new
requirements, in both functionality and performance, and this is given
a high priority by the HEP community.

@@

One new model of data analysis, developed outside of HEP, maintains the
concept of sequential reduction, but mixes interactivity with
batch processing. These exploit new cluster management systems, most
notably Apache Spark~\cite{Spark}, which uses open-source tools contributed both by
industry and the data-science community. Other products implementing the
same analysis concepts and workflows are emerging, such as
TensorFlow~\cite{abadi2016tensorflow}, Dask~\cite{rocklin2015dask}, Pachyderm~\cite{thepachydermteam}, Blaze~\cite{wiebe2014blaze}, Parsl~\cite{babuji_yadu_2017_853492}, and Thrill~\cite{bingmann2016thrill}.
This approach can complement the present and widely adopted Grid processing of datasets. It may
potentially simplify the access to data and the expression of
parallelism, thereby improving the exploitation of cluster resources.

>>>>

\begin{figure*}
\vspace*{0.3cm}
    \centering
    \includegraphics[width=0.94\textwidth]{10yr-PLAN-9Feb-2017-V7a.pdf}
    \caption{Run schedule for the Fermilab facility until 2026~\cite{fnal-schedule}.}
    \label{fig:fnal-if}
\end{figure*}

@@

While the density of transistors on silicon continues to increase
following Moore's Law (albeit more slowly than in the past), power
density constraints have limited the clock speed of processors for more
than a decade~\cite{NAP12980}. This has effectively stalled any progress in the
processing capacity of a single CPU core. Instead, increases in
potential processing capacity come from increases in the core count of
CPUs and wide CPU registers. Alternative processing architectures have
become more commonplace. These range from the many-core architecture
based on standard x86\_64 cores to numerous alternatives such as GPUs.
For GPUs, the processing model is very different~\cite{Cook:2012:CPD:2430671}, allowing a much
greater fraction of the die to be dedicated to arithmetic calculations,
but at a price in programming difficulty and memory handling for the
developer that tends to be specific to each processor generation.
Further developments may even see the use of FPGAs for more
general-purpose tasks. Fully exploiting these evolutions requires a
shift in programming model to one based on \emph{concurrency}.

@@

Particle physics is no longer alone in facing these massive data
challenges. Experiments in other fields, from astronomy~\cite{10.1371/journal.pbio.1002195} 
to genomics~\cite{ijms18020412},
will produce huge amounts of data in the future, and will need to
overcome the same challenges that we face, i.e., massive data handling and
efficient scientific programming. Establishing links with these fields
has already started. Additionally, interest from the computing science
community in solving these data challenges exists, and mutually
beneficial relationships would be possible where there are genuine
research problems that are of academic interest to that community and
provide practical solutions to ours. The efficient processing of massive
data volumes is also a challenge faced by industry, in particular the
internet economy, which developed novel and major new technologies under
the banner of \emph{Big Data} that may be applicable to our use cases
\cite{47224,Shanahan:2015:LSD:2783258.2789993,Armbrust:2015:SSR:2723372.2742797}.

@@

\paragraph{Fast Simulation} The work on Fast Simulation is also
accelerating with the objective of producing a flexible framework that
permits Full and Fast simulation to be combined for different particles
in the same event. Various approaches to Fast Simulation are being tried
all with the same goal of saving computing time, under the assumption
that it is possible to improve time performance without an unacceptable
loss of physics accuracy. There has recently been a great deal of
interest in the use of Machine Learning in Fast Simulation, most of
which has focused on the use of multi-objective regression and
generative adversarial networks (GANs)~\cite{NIPS2014_5423}. Since use of GANs allows for
non-parametric learning in cases such as calorimetric shower
fluctuations, it is a promising avenue for generating non-Gaussian and
highly correlated physical effects. This is an obvious area for future
expansion and development, as it is currently in its infancy.

@@

Over the past 20 years the HEP community has developed and gravitated
around a single analysis ecosystem based on ROOT.

ROOT is a general-purpose object
oriented framework that addresses the selection, integration,
development, and support of a number of foundation and utility class
libraries that can be used as a basis for developing HEP application
codes. The added value to the HEP community is that it provides an
integrated and validated toolkit, and its use encompasses the full event
processing chain; it has a major impact on the way HEP analysis is
performed. This lowers the hurdle to start an analysis, enabling the
community to communicate using a common analysis language, as well as
making common improvements as additions to the toolkit quickly become
available. The ongoing ROOT programme of work addresses important new
requirements, in both functionality and performance, and this is given
a high priority by the HEP community.

@@

One new model of data analysis, developed outside of HEP, maintains the
concept of sequential reduction, but mixes interactivity with
batch processing. These exploit new cluster management systems, most

notably Apache Spark~\cite{Spark,Armbrust:2015:SSR:2723372.2742797}, which uses open-source tools contributed both by
industry and the data-science community. Other products implementing the
same analysis concepts and workflows are emerging, such as TensorFlow~\cite{tensorflow2015-whitepaper},
Dask~\cite{rocklin2015dask,dask_ref}, Pachyderm~\cite{thepachydermteam}, 
Blaze~\cite{wiebe2014blaze}, Parsl~\cite{babuji_yadu_2017_853492}, 
and Thrill~\cite{bingmann2016thrill}. This approach can complement
the present and widely adopted Grid processing of datasets. It may

potentially simplify the access to data and the expression of
parallelism, thereby improving the exploitation of cluster resources.
